---
title: "Hands-on Exercise 6: Spatial Weights and Applications"
date: "12 Feb 2023"
date-modified: "`r Sys.Date()`"
format: html
execute: 
  message: true
  warning: false
editor: visual
---

# Getting started

```{r}
pacman:: p_load(sf, spdep, tmap, tidyverse, knitr)
```

# Study area

## Import shapefile of Hunan

```{r}
hunan <- st_read(dsn="data/geospatial",
                     layer="Hunan") 
```

## Import csv of Hunan population

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

## Relational join

```{r}
hunan <- left_join(hunan,hunan2012)%>%
  select(1:4, 7, 15)
```

# Regional development - GPD

```{r}
basemap <- tm_shape(hunan) +
  tm_polygons() +
  tm_text("NAME_3", size=0.5)

gdppc <- qtm(hunan, "GDPPC")
tmap_arrange(basemap, gdppc, asp=1, ncol=2)
```

# Contiguity Spatial Weights

## QUEEN contiguity based neighbours

```{r}
wm_q <- poly2nb(hunan, queen=TRUE)
summary(wm_q)
```

The summary report above shows that there are 88 area units in Hunan. The most connected area unit has 11 neighbours. There are two area units with only one heighbours.

For each polygon in our polygon object, wm_q lists all neighboring polygons. For example, to see the neighbors for the first polygon in the object, type:

```{r}
wm_q[[1]]
```

Polygon 1 has 5 neighbors. The numbers represent the polygon IDs as stored in hunan SpatialPolygonsDataFrame class.

We can retrive the county name of Polygon ID=1 by using the code chunk below:

```{r}
hunan$County[1]
```

The output reveals that Polygon ID=1 is Anxiang county.

To reveal the county names of the five neighboring polygons, the code chunk will be used:

```{r}
hunan$NAME_3[c(2,3,4,57,85)]
```

We can retrieve the GDPPC of these five countries by using the code chunk below.

```{r}
nb1 <- wm_q[[1]]
nb1 <- hunan$GDPPC[nb1]
nb1
```

The printed output above shows that the GDPPC of the five nearest neighbours based on Queen's method are 20981, 34592, 24473, 21311 and 22879 respectively.

You can display the complete weight matrix by using str().

```{r}
str(wm_q)
```

## ROOK contiguity based neighbours

```{r}
wm_r <- poly2nb(hunan, queen=FALSE)
summary(wm_r)
```

The summary report above shows that there are 88 area units in Hunan. The most connect area unit has 10 neighbours. There are two area units with only one neighbour.

## Visualising contiguity weights

A connectivity graph takes a point and displays a line to each neighboring point. We are working with polygons at the moment, so we will need to get points in order to make our connectivity graphs. The most commonly used method for this will be polygon centroids. We will calculate these in the sf package before moving onto the graphs (getting Latitude and Longitude of Polygon Centroids).

We will need points to associate with each polygon before we can make our connectivity graph. It will be a little more complicated than just running st_centroid on the sf object: us.bound. We need the coordinates in a separate data frame for this to work.

To do this we will use a mapping function. The mapping function applies a given function to each element of a vector and returns a vector of the same length. Our input vector will be the geometry column of us.bound. Our function will be st_centroid. We will be using map_dbl variation of map from the purrr package. For more documentation, check out map documentation.

To get our longitude values we map the st_centroid function over the geometry column of us.bound and access the longitude value through double bracket notation \[\[\]\] and 1. This allows us to get only the longitude, which is the first value in each centroid.

```{r}
longitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[1]])
```

Do the same for latitude, but access the second value per centroid \[\[2\]\].

```{r}
latitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[2]])
```

```{r}
coords <- cbind(longitude, latitude)
```

```{r}
head(coords)
```

## Plotting QUEEN contiguity based neighbours map

```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red")
```

## Plotting ROOK contiguity based neighbours map

```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

## Plotting both QUEEN and ROOK

```{r}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red", main="Queen Contiguity")
plot(hunan$geometry, border="lightgrey")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red", main="Rook Contiguity")
```

# Distance based neighbours

In this section, you will learn how to derive distance-based weight matrices by using dnearneigh() of spdep package.

The function identifies neighbours of region points by Euclidean distance with a distance band with lower d1= and upper d2= bounds controlled by the `bounds=` argument. If unprojected coordinates are used and either specified in the coordinates object x or with x as a two column matrix and `longlat=TRUE`, great circle distances in km will be calculated assuming the WGS84 reference ellipsoid.

## Cut-off distance

Firstly, we need to determine the upper limit for distance band by using the steps below:

-   Return a matrix with the indices of points belonging to the set of the k nearest neighbours of each other by using `knearneigh()` of spdep.
-   Convert the knn object returned by `knearneigh()` into a neighbours list of class nb with a list of integer vectors containing neighbour region number ids by using `knn2nb()`.
-   Return the length of neighbour relationship edges by using `nbdists()` of spdep. The function returns in the units of the coordinates if the coordinates are projected, in km otherwise.
-   Remove the list structure of the returned object by using `unlist()`.

```{r}
#coords <- coordinates(hunan)
k1 <- knn2nb(knearneigh(coords))
k1dists <- unlist(nbdists(k1, coords, longlat = TRUE))
summary(k1dists)
```

The summary report shows that the largest first nearest neighbour distance is 61.79 km, so using this as the upper threshold gives certainty that all units will have at least one neighbour.

## Computing fixed distance weight matric

```{r}
wm_d62 <- dnearneigh(coords, 0, 62, longlat = TRUE)
wm_d62
```

```{r}
str(wm_d62)
```

Another way of displaying structure of weight matrix:

```{r}
table(hunan$County, card(wm_d62))
```

```{r}
n_comp <- n.comp.nb(wm_d62)
n_comp$nc
```

```{r}
table(n_comp$comp.id)
```



## Computing fixed distance weight matrix
```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_d62, coords, add=TRUE)
plot(k1, coords, add=TRUE, col="red", length=0.08)
```
The red lines show the links of 1st nearest neighbours and the black lines show the links of neighbours within the cut-off distance of 62km.

Alternatively, we can plot both of them next to each other by using the code chunk below.

```{r}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey")
plot(k1, coords, add=TRUE, col="red", length=0.08, main="1st nearest neighbours")
plot(hunan$geometry, border="lightgrey")
plot(wm_d62, coords, add=TRUE, pch = 19, cex = 0.6, main="Distance link")
```

## Computing Adaptive distance weight matrix
One of the characteristics of fixed distance weight matrix is that more densely settled areas (usually the urban areas) tend to have more neighbours and the less densely settled areas (usually the rural counties) tend to have lesser neighbours. Having many neighbours smoothes the neighbour relationship across more neighbours.

It is possible to control the numbers of neighbours directly using k-nearest neighbours, either accepting asymmetric neighbours or imposing symmetry as shown in the code chunk below.

```{r}
knn6 <- knn2nb(knearneigh(coords, k=6))
knn6
```
View contents of matrix:
```{r}
str(knn6)
```

Each county only has exactly 6 neighbours!

## Distance based neighbours
```{r}
plot(hunan$geometry, border="lightgrey")
plot(knn6, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

# Weights based on Inverse Distance method (IDW)

```{r}
dist <- nbdists(wm_q, coords, longlat = TRUE)
ids <- lapply(dist, function(x) 1/(x))
ids
```

## Row standardised weights matrix
Next, we need to assign weights to each neighboring polygon. In our case, each neighboring polygon will be assigned equal weight (style=“W”). This is accomplished by assigning the fraction 1/(#ofneighbors) to each neighboring county then summing the weighted income values. While this is the most intuitive way to summaries the neighbors’ values it has one drawback in that polygons along the edges of the study area will base their lagged values on fewer polygons thus potentially over- or under-estimating the true nature of the spatial autocorrelation in the data. For this example, we’ll stick with the style=“W” option for simplicity’s sake but note that other more robust options are available, notably style=“B”.


```{r}
rswm_q <- nb2listw(wm_q, style="W", zero.policy = TRUE)
rswm_q
```

The zero.policy=TRUE option allows for lists of non-neighbors. This should be used with caution since the user may not be aware of missing neighbors in their dataset however, a zero.policy of FALSE would return an error.

To see the weight of the first polygon’s eight neighbors type:

```{r}
rswm_q$weights[10]
```
Each neighbor is assigned a 0.125 of the total weight. This means that when R computes the average neighboring income values, each neighbor’s income will be multiplied by 0.2 before being tallied.

Using the same method, we can also derive a row standardised distance weight matrix by using the code chunk below.


```{r}
rswm_ids <- nb2listw(wm_q, glist=ids, style="B", zero.policy=TRUE)
rswm_ids
```

```{r}
rswm_ids$weights[1]
```

```{r}
summary(unlist(rswm_ids$weights))
```

# Application of spatial weight matrix

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```
