---
title: "Take Home Exercise 3"
date: "__ 2023"
date-modified: "`r Sys.Date()`"
format: html
execute: 
  message: true
  warning: false
editor: visual
---

```{r}
pacman::p_load(tidygeocoder, sf, tidyverse, tmap, jsonlite )
```


# aspatial data
```{r}
resale_all <- read_csv("data/aspatial/resale-flat-prices-based-on-registration-date-from-jan-2017-onwards.csv")
```

```{r}
glimpse(resale_all)
```
::: 
if (grepl('2017', resale_all$month)==TRUE & grepl('2018', resale_all$month)==TRUE & grepl('2019', resale_all$month)==TRUE ){
  resale17 <- resale_all %>% filter(resale_all, month!='2017')
:::

using grepl() function [https://www.programiz.com/r/examples/check-if-character-present-in-string]
```{r}
resale2122 <- filter(resale_all, grepl('2021', month)|grepl('2022', month))
```

```{r}
resale23_test <- filter(resale_all, grepl('2023', month))
```

```{r}
n_distinct(resale2122$flat_type)
```
```{r}
unique(resale2122$flat_type)
```


```{r}
nrow(resale2122[resale2122$flat_type == '2 ROOM',])
```
```{r}
nrow(resale_all[resale_all$flat_type == '3 ROOM',])
```
```{r}
nrow(resale_all[resale_all$flat_type == '4 ROOM',])
```
```{r}
nrow(resale_all[resale_all$flat_type == '5 ROOM',])
```
```{r}
resale2122_5r <- filter(resale2122, flat_type == '5 ROOM')
```

geocode resale data using tidygeocoder package
https://cran.r-project.org/web/packages/tidygeocoder/readme/README.html 
```{r}
geocoded <- resale2122_5r %>%
  geocode(street_name, method = 'osm', lat = latitude , long = longitude)
geocoded
```

```{r}
sum(is.na(geocoded$longitude))
```


geocoding using onemap API
```{r}
resale2122_5r$street_name <- gsub("ST\\.", "SAINT", resale2122_5r$street_name)

```

```{r}
library(httr)
geocode_function <- function(block, street_name) {
  base_url <- "https://developers.onemap.sg/commonapi/search"
  address <- paste(block, street_name, sep = " ")
  query <- list("searchVal" = address, 
                "returnGeom" = "Y",
                "getAddrDetails" = "N",
                "pageNum" = "1")
  
  res <- GET(base_url, query = query)
  restext<-content(res, as="text")
  
  output <- fromJSON(restext)  %>% 
    as.data.frame %>%
    select(results.LATITUDE, results.LONGITUDE)

  return(output)
}
```


```{r}
resale2122_5r$LATITUDE <- 0
resale2122_5r$LONGITUDE <- 0

for (i in 1:nrow(resale2122_5r)){
  temp_output <- geocode_function(resale2122_5r[i, 4], resale2122_5r[i, 5])
  
  resale2122_5r$LATITUDE[i] <- temp_output$results.LATITUDE
  resale2122_5r$LONGITUDE[i] <- temp_output$results.LONGITUDE
}
```

```{r}
sum(is.na(resale2122_5r$LATITUDE))

```

```{r}
sum(is.na(resale2122_5r$LONGITUDE))
```

```{r}

```

# geospatial data

locational factors

- MRT stations [lta datamall]
- bus stops [lta datamall]
- supermarkets 
- parks
- hawker centre
- eldercare services
- primary schools within 1km
- good primary schools from <a href="https://www.salary.sg/2022/best-primary-schools-2022-by-popularity/">link</a> top 20 schools

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```


```{r}

```

```{r}

```

```{r}

```

```{r}

```

